### 2025-01-09 PbD/exercises.tex
Added some more exercises.
Added exercises into latex.

>>>>

\section{Assignments}
\begin{exercise}{Protection Goals}
  In IT security, the protection goals of confidentiality, availability, and integrity are defined to describe what objectives an attacker aims for or what a technology is designed to protect. Provide an appropriate technique for each of these protection goals that can be used to implement or address the respective goal.

  \begin{solution}
    \begin{itemize}
      \item \textbf{Confidentiality}: Encryption is used to protect the confidentiality of data.
      \item \textbf{Availability}: Redundancy is used to ensure the availability of services.
      \item \textbf{Integrity}: Hash functions are used to ensure the integrity of data.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Pseudonyms}
  Pseudonyms are fundamental building blocks of secure and privacy-friendly systems.
  \begin{enumerate}
    \item Briefly describe the general purpose of pseudonyms.
    \item What is the difference between pseudonymity and anonymity?
    \item Name three applications in which pseudonyms are used. Specify the type of pseudonym used in each case. Explain how pseudonyms are employed in each application.
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item The general purpose of pseudonyms is to protect the identity of users.
      \item The difference between pseudonymity and anonymity is that pseudonymity uses a pseudonym to protect the identity of users, while anonymity does not require a pseudonym.
      \item Three applications in which pseudonyms are used are:
        \begin{itemize}
          \item E-Mail: Pseudonyms are used to protect the identity of users.
          \item Social Media: Pseudonyms are used to protect the identity of users.
          \item Online Shopping: Pseudonyms are used to protect the identity of users.
        \end{itemize}
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Indeterministic Encryption}
  \begin{enumerate}
    \item What does indeterministic encryption mean?
    \item How do you implement indeterministic encryption in an asymmetric cryptosystem?
    \item What changes compared to b) when using a symmetric cryptosystem?
    \item You are tasked with implementing indeterministic encryption. The algorithms RSA and ElGamal are available. How would you proceed with each algorithm?
    \item Which attack can be prevented by using indeterministic encryption?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item Indeterministic encryption means that the same plaintext is encrypted to different ciphertexts each time.
      \item Indeterministic encryption in an asymmetric cryptosystem is implemented by adding randomness to the encryption process.
      \item In a symmetric cryptosystem, indeterministic encryption is implemented by adding randomness to the encryption process.
      \item To implement indeterministic encryption with RSA, random padding is added to the plaintext before encryption. With ElGamal, a random number is used to generate the encryption key.
      \item Indeterministic encryption prevents chosen-ciphertext attacks.
    \end{enumerate}
  \end{solution}
\end{exercise}

% Aufgabe 4 Unobservable retrieval of information
% Given is a Blind-Message-Service with three servers S1, S2, and S3. The following datasets are stored in the
% Blind-Message-Service: D0 = 1011, D1 = 0101, D2 = 1100, D3 = 0011, D4 = 0000.
% a) You want to retrieve dataset D2 without the servers knowing which dataset you are interested in. What
% messages are transmitted? How do you obtain the requested dataset? Sketch the computational steps.
% b) You want to ensure, as in task a), that during a query it remains unknown which dataset the client is
% interested in. However, instead of the Blind-Message-Service, only a single high-bandwidth server is
% available. The database, however, is relatively small. Which method do you apply?
\begin{exercise}{Unobservable Retrieval of Information}
  Given is a Blind-Message-Service with three servers S1, S2, and S3. The following datasets are stored in the Blind-Message-Service: D0 = 1011, D1 = 0101, D2 = 1100, D3 = 0011, D4 = 0000.
  \begin{enumerate}
    \item You want to retrieve dataset D2 without the servers knowing which dataset you are interested in. What messages are transmitted? How do you obtain the requested dataset? Sketch the computational steps.
    \item You want to ensure, as in task a), that during a query it remains unknown which dataset the client is interested in. However, instead of the Blind-Message-Service, only a single high-bandwidth server is available. The database, however, is relatively small. Which method do you apply?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item To retrieve dataset D2 without the servers knowing which dataset you are interested in, you send a request for all datasets and then filter out the requested dataset. The computational steps are as follows:
        \begin{itemize}
          \item Send a request for all datasets.
          \item Receive the datasets from the servers.
          \item Filter out the requested dataset D2.
        \end{itemize}
      \item To ensure that the client's interest remains unknown, you use the Broadcast method.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Tracking on the Application Layer}
  \begin{enumerate}
    \item Name two security goals that are violated during tracking at the application layer.
    \item What is the difference between tracking at the application layer and tracking at the network layer?
    \item Name three fundamental mechanisms to prevent or make tracking at the application layer more difficult. Explain for each mechanism how it hinders tracking, against which type of attacker, and provide a specific example where the mechanism is implemented.
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item Two security goals that are violated during tracking at the application layer are confidentiality and anonymity.
      \item The difference between tracking at the application layer and tracking at the network layer is that tracking at the application layer uses browser data, while tracking at the network layer uses IP data.
      \item Three fundamental mechanisms to prevent or make tracking at the application layer more difficult are:
        \begin{itemize}
          \item Block cookies: Prevents tracking by websites. Against advertisers. Example: Browser settings.
          \item Use VPNs: Prevents tracking by ISPs. Against ISPs. Example: NordVPN.
          \item Use privacy plugins: Prevents tracking by websites. Against advertisers. Example: uBlock Origin.
        \end{itemize}
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Surveillance on the Internet}
  \begin{enumerate}
    \item Explain a possible approach by which a powerful attacker (e.g., an intelligence agency) can determine the IP addresses of users accessing a specific website through a mix-based anonymization system (e.g., Tor).
    \item Explain a possible approach by which a powerful attacker (e.g., an intelligence agency) can determine which websites a specific user has accessed via a mix-based anonymization system. Note: Describe a different method from the one in task a), even if your method in a) could also be applied to task b).
    \item Explain three techniques by which a web server can recognize a user during their second (subsequent) visit.
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item A powerful attacker can determine the IP addresses of users accessing a specific website through a mix-based anonymization system by using traffic analysis and traffic correlation.
      \item A powerful attacker can determine which websites a specific user has accessed via a mix-based anonymization system by using metadata analysis, such as HTTP headers or specific request characteristics.
      \item Three techniques by which a web server can recognize a user during their second visit are:
        \begin{itemize}
          \item Cookies: Store user data in the browser. Against websites. Example: Google Analytics.
          \item Browser fingerprinting: Collect browser data. Against websites. Example: Panopticlick.
          \item Tracking pixels: Track user behavior. Against advertisers. Example: Facebook Pixel.
        \end{itemize}
    \end{enumerate}
  \end{solution}
\end{exercise}
